---
title: "Project 2"
author: "Rohan Prabhune, Naman Goel"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---
```{r Temp2,echo=FALSE,eval=TRUE,message=FALSE,warning=FALSE}
wrapper <- function(x, ...) 
{
  paste(strwrap(x, ...), collapse = "\n")
}
#knitr::opts_chunk$set(fig.width=8, fig.height=7) 
knitr::opts_chunk$set(fig.align = 'center') 
```
# Required packages list
```{r 1_1,echo=TRUE,eval=TRUE,message=FALSE,warning=FALSE}
library(httr)
library(jsonlite)
library(tidyverse)
library(knitr)
library(ggplot2)
library(kableExtra)
library(patchwork)
library(rmarkdown)
```
<!--*************************************************************************-->
# Functions to query endpoints
## Endpoint 1 - Four modifications
```{r 2_1,echo=TRUE,eval=TRUE,message=FALSE,warning=FALSE}
get_stocks_agg <- function(ticker,company_name,start_date="2022-01-01",
                           end_date="2022-08-31",limit=50){
  url=paste0("https://api.polygon.io/v2/aggs/ticker/",ticker,
             "/range/1/day/",start_date,"/",end_date,
             "?adjusted=true&sort=asc&limit=",limit,"&apiKey=EdkA7_m2JhjS5POrGuXJbVlA4AjSl_4F")
  response_obj <- GET(url)
  parsed <- fromJSON(rawToChar(response_obj$content))
  df <- as_tibble(parsed$results)
  df <- df %>% rename(close_price=c,highest_price=h,lowest_price=l,num_transactions=n,
                      open_price=o,timestamp=t,vol=v,weighted_avg_price=vw) 
  df$ticker <- rep(ticker,limit)
  df$company_name <- rep(company_name,limit)
  df$start_date <- start_date
  df$end_date <- end_date
  df <- df %>% select(ticker,company_name,everything())
  return(df)
}
#TODO: Convert timestamp
```

## Endpoint 2 - Two modifications
Searches only stocks.Gives the first ticker if the name is not complete. Limit is 1000 and its sorted in ascending order.
```{r 2_2,echo=TRUE,eval=TRUE,message=FALSE,warning=FALSE}
get_ticker_symbol <- function(name){
  url=paste0("https://api.polygon.io/v3/reference/tickers?market=stocks&search=",name,"&active=true&sort=ticker&order=asc&limit=1000&apiKey=EdkA7_m2JhjS5POrGuXJbVlA4AjSl_4F")
  response_obj <- GET(url)
  parsed<- fromJSON(rawToChar(response_obj$content))
  df <- as_tibble(parsed$results)
  return(list(df$ticker[[1]],df$name[[1]]))
}
```

```{r 2_3,echo=TRUE,eval=TRUE,message=FALSE,warning=FALSE}
get_ticker_info <- function(market){
  url=paste0("https://api.polygon.io/v3/reference/tickers?market=",market,"&active=true&sort=ticker&order=asc&limit=1000&apiKey=EdkA7_m2JhjS5POrGuXJbVlA4AjSl_4F")
  response_obj <- GET(url)
  parsed<- fromJSON(rawToChar(response_obj$content))
  df <- as_tibble(parsed$results) %>% select(ticker,name,primary_exchange,type)
  return(df)
}
```

## Endpoint 3 - One modification
```{r 2_4,echo=TRUE,eval=TRUE,message=FALSE,warning=FALSE}
get_grouped_daily <- function(date="2020-10-14"){
  url=paste0("https://api.polygon.io/v2/aggs/grouped/locale/us/market/stocks/",date,"?adjusted=true&include_otc=true&apiKey=EdkA7_m2JhjS5POrGuXJbVlA4AjSl_4F")
  response_obj <- GET(url)
  parsed <- fromJSON(rawToChar(response_obj$content))
  df <- as_tibble(parsed$results)
  df <- df %>% rename(Ticker=T,volume=v,weighted_avg_price=vw,open_price=o,
                      close_price=c,highest_price=h,lowest_price=l,
                      num_transactions=n,timestamp=t)
  df$date <- date
  return(df)
}
```

## Endpoint 4 
```{r 2_5,echo=TRUE,eval=TRUE,message=FALSE,warning=FALSE}
get_ticker_type_details <- function(){
  response_obj <- GET("https://api.polygon.io/v3/reference/tickers/types?apiKey=EdkA7_m2JhjS5POrGuXJbVlA4AjSl_4F")
  parsed <- fromJSON(rawToChar(response_obj$content))
  df <- as_tibble(parsed$results) 
  return(df)
}
```

## Endpoint 5 
```{r 2_6,echo=TRUE,eval=TRUE,message=FALSE,warning=FALSE}
get_exchanges_details <- function(){
  response_obj <- GET("https://api.polygon.io/v3/reference/exchanges?asset_class=stocks&apiKey=EdkA7_m2JhjS5POrGuXJbVlA4AjSl_4F")
  parsed <- fromJSON(rawToChar(response_obj$content))
  df <- as_tibble(parsed$results)  
  return(df)
}
```
<!--*************************************************************************-->
# Creation of relevant new variables
```{r 3_1,echo=TRUE,eval=TRUE,message=FALSE,warning=FALSE}
df_grouped <- get_grouped_daily("2020-11-16")
df_grouped <- df_grouped %>% mutate(percent_change = round(((close_price-open_price)/open_price)*100,2)) %>% arrange(desc(percent_change))
df_grouped$percent_change_chr <- paste(as.character(df_grouped$percent_change),'%')
df_grouped %>% select(Ticker,open_price,close_price,percent_change_chr,everything())
```

<!--*************************************************************************-->
# Combining data from API Calls
```{r 4_1,echo=TRUE,eval=TRUE,message=FALSE,warning=FALSE}
ticker_symbol1 <- get_ticker_symbol(name="Apple")
df1 <- get_stocks_agg(ticker_symbol1[[1]],ticker_symbol1[[2]],start_date="2022-01-01",end_date="2022-08-31")
df1

ticker_symbol2 <- get_ticker_symbol(name="Tesla")
df2 <- get_stocks_agg(ticker_symbol2[[1]],ticker_symbol2[[2]],start_date="2022-01-01",end_date="2022-08-31")
df2

ticker_symbol3 <- get_ticker_symbol(name="Nvidia")
df3 <- get_stocks_agg(ticker_symbol3[[1]],ticker_symbol3[[2]],start_date="2022-01-01",end_date="2022-08-31")
df3
```

```{r 4_2,echo=TRUE,eval=TRUE,message=FALSE,warning=FALSE}
df_combined <- bind_rows(df1, df2, df3)
ggplot(df_combined,aes(x=timestamp,y=close_price)) + geom_line(aes(color=company_name),size=1) + labs(x="Time",y="Closing price",title="Closing stock price over time for Apple, Nvidia and Tesla") +  scale_color_discrete(name = "Company Name")+theme(plot.title = element_text(hjust = 0.5))
```

<!--*************************************************************************-->
# Contingency tables
```{r 4_3,echo=TRUE,eval=TRUE,message=FALSE,warning=FALSE}
df_info <- get_ticker_info("stocks")

tab1 <- table(df_info$type,dnn=c("Ticker Types"))
tab2 <- table(df_info$type,df_info$primary_exchange,dnn=c("Ticker Types","Exchanges"))

tab1 %>%
  kbl(caption="Table for Ticker Types") %>%
  kable_classic(full_width = F)
tab2 %>%
  kbl(caption="Table for Ticker Types and Exchanges") %>%
  kable_classic(full_width = F)
```

<!--*************************************************************************-->
# Numerical summaries
```{r 5_1,echo=TRUE,eval=TRUE,message=FALSE,warning=FALSE}
df_combined <- bind_rows(df1, df2, df3)
start_date = as.character(unique(df_combined$start_date))
end_date = as.character(unique(df_combined$end_date))
# Open Price
df_combined_open <- df_combined %>% 
  group_by(company_name) %>% 
  summarise("Min." = min(open_price),
            "1st Quartile" = quantile(open_price,0.25),
            "Median." = median(open_price),
            "Mean."=mean(open_price),
            "3rd Quartile" = quantile(open_price,0.75),
            "Max."= max(open_price),
            "Std. Dev." = sd(open_price))

# Close Price
df_combined_close <- df_combined %>% 
  group_by(company_name) %>% 
  summarise("Min." = min(close_price),
            "1st Quartile" = quantile(close_price,0.25),
            "Median." = median(close_price),
            "Mean."=mean(close_price),
            "3rd Quartile" = quantile(close_price,0.75),
            "Max."= max(close_price),
            "Std. Dev." = sd(close_price))

df_combined_open %>%
  kbl(caption=paste0("Summary Statistics for Open Price per Company between ",start_date," and ",end_date)) %>%
  kable_classic()

df_combined_close %>%
  kbl(caption=paste0("Summary Statistics for Close Price per Company between ",start_date," and ",end_date)) %>%
  kable_classic()
```

<!--*************************************************************************-->
# Plots
Did not map the ticker symbol to the company name because there is a max limit of 1000 on the "Ticker" reference end point. So we do not get all the data.

## Plotting new created variable
```{r 6_1,echo=TRUE,eval=TRUE,message=FALSE,warning=FALSE}
#Top-10 increases
df_top10 <- head(df_grouped, 10)
df_top10$Ticker <- factor(df_top10$Ticker, levels=df_top10$Ticker[order(-df_top10$percent_change)])

# Plot
date <- unique(df_top10$date)
ggplot(df_top10, aes(x=Ticker, y=percent_change)) + 
geom_col(width=0.3, color='steelblue', fill='steelblue') + 
theme(axis.text.x=element_text(angle=90), text=element_text(size=12), plot.title = element_text(hjust = 0.5)) + labs(y="Percent increase", x ="Stock ticker",title = paste0("Highest stock price increase on ",date)) + geom_text(aes(label = percent_change_chr), vjust = -0.5, size=3)

#Top-10 decreases
df_bottom10 <- tail(df_grouped, 10)
df_bottom10$percent_change <- abs(df_bottom10$percent_change)
df_bottom10$Ticker <- factor(df_bottom10$Ticker, levels=df_bottom10$Ticker[order(-df_bottom10$percent_change)])

# Plot
date = unique(df_bottom10$date)
ggplot(df_bottom10, aes(x=Ticker, y=percent_change)) + 
geom_col(width=0.3, color='red', fill='red') + 
theme(axis.text.x = element_text(angle=90),text = element_text(size=12),plot.title = element_text(hjust = 0.5)) + labs(y="Percent decrease", x ="Stock ticker", title = paste0("Highest stock price decrease on ",date)) + geom_text(aes(label=percent_change_chr), vjust =-0.5, size=3)
```

## Box plots for highest and lowest price per company in a given time period
```{r 6_2,echo=TRUE,eval=TRUE,message=FALSE,warning=FALSE}
start_date = as.character(unique(df_combined$start_date))
end_date = as.character(unique(df_combined$end_date))

ggplot(df_combined, aes(x=company_name, y=highest_price)) + geom_boxplot(color="blue",fill="grey") + labs(y="Highest price", x ="Company Name", title=paste0("Boxplot for highest stock price between ",start_date," and ",end_date)) 

ggplot(df_combined, aes(x=company_name, y=lowest_price)) + geom_boxplot(color="red",fill="grey") + labs(y="Lowest price", x ="Company Name", title=paste0("Boxplot for lowest stock price between ",start_date," and ",end_date)) 
```

## Histogram 
```{r 6_3,echo=TRUE,eval=TRUE,message=FALSE,warning=FALSE}
start_date = as.character(unique(df_combined$start_date))
end_date = as.character(unique(df_combined$end_date))

my_title <- paste0("Histogram of weighted average price per company between ",start_date," and ",end_date)
ggplot(df_combined, aes(x=weighted_avg_price)) + geom_histogram(aes(fill=company_name),binwidth=8) + labs(x ="Weighted average price") + scale_fill_discrete(name = "Company Name") + ggtitle(wrapper(my_title, width=80))
```

## Bar plot 
```{r 6_4,echo=TRUE,eval=TRUE,message=FALSE,warning=FALSE}
ggplot(df_info, aes(x=type)) + geom_bar(fill="steelblue") + labs(x ="Type of tickers",title="Bar plot for number of stock tickers for each type") + theme(text=element_text(size=12), plot.title = element_text(hjust = 0.5)) + geom_text(aes(label = ..count..), stat = "count", vjust = -0.5)
```

## Scatter plot  
The scatter plot has been used to plot the percentage change in returns for the stocks Apple, Tesla and Nvidia for a date range and can be differentiated based on the different colours. From the plot, we can see how varied returns have been on a daily basis but if we look carefully we can see that there is a certain correlation in the returns and on most days the movement for all 3 technology stocks is in the same direction of either gaining or losing
```{r 6_5, echo=TRUE,eval=TRUE,message=FALSE,warning=FALSE}
df_comb <- df_combined %>% mutate(percent_change = round(((close_price-open_price)/open_price)*100,2))
ggplot(df_comb, aes(x=timestamp, y=percent_change, color=ticker)) + geom_point()
```
